\MakeUppercase{Layered $H$-Partitions of $k$-Planar Graphs} \thanks{This work was partly funded by NSERC and MRI.}

### 2019-05-16 k-planar.tex
Almost done with k^3 proof

    \item $\ell\le k+1$.
  \end{compactenum}
  We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists an $\delta$-walk.

  In the following, when $i$ is an integer, we will use $F_i$, $Y_i$, and so on as shorthands for $F_{a_i}$, $Y_{a_i}$, and so on, respectively.

  To begin, let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  Observe that $F_{d}=F_{d_0}$ is contained in the interior of $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.   Refer to \figref{d0d1}.  We claim that 
  \[  
     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace .
  \]
  \begin{figure}
    \begin{center}
      \includegraphics{figs/d0d1-5}
    \end{center}
    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.}
    \figlabel{d0d1}
  \end{figure}
  To prove this claim, consider the graph $I$ with vertex set $V(I)=\{d_1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.  Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.

  The graph $I$ has the following properties:
  \begin{enumerate}
    \item For each $i\in V(I)$, $i$ has at most 2 neighbours among $d_1,\ldots,i-1$.  If $i$ has two such neighbours, then one of them is $i-1$.

    \item For each $ij\in E(I)$ with $j<i$, and each $i'\in\{j+2,\ldots,i\}$, the vertex $i'$ has at most 2 neighbours and these are contained in $\{i'-1,j\}$.
  \end{enumerate}
  These two properties imply that $I$ has a very simple structure.   We call $j\in\{d_1,\ldots,d_0\}$ a \emph{left endpoint} if there exists an edge $ij\in E(I)$ with $j < i - 1$.  Let $j_1>j_2>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.  Then 
  \[
      E(I) \subseteq \{ (i,i+1): i\in\{d_1,\ldots,d_0-1\} \cup 
         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1} \}\} \enspace . 
  \]
  (See the lower half of \figref{d0d1}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.  Let $X_t=\{j_t,\ldots,j_{t+1}\}$ and observe that the shortest path from $d_0$ to some value $i\in X_t$ that does not contain any $i' <i$ goes through $j_{t}$.  Therefore, $X_t$ contains at most $k+2-t$ values of $\delta$ such that there exists a $\delta$-walk.  Therefore, 
  \[  
     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+2}{2} \enspace .
  \]

>>>>

    \item $\ell\le k+1$.
  \end{compactenum}
  We call any walk with properties (i)--(iv) a \emph{$\delta$-walk}. Therefore, to upper-bound $|B_x|$, it suffices to upper-bound the number of $\delta\in\{0,\ldots,d-1\}$ such that there exists a $\delta$-walk.

  In order to avoid excessive levels of subscripts the remainder of this proof will slightly abuse notation in the following way: For $i\in\{0,\ldots,d\}$, we use $F_i$, $N_i$, and $Y_i$, as shorthands for $F_{a_i}$, $Y_{a_i}$, respectively.

  To begin, let $d_0=d$, and $d_1=\min\{i: V(F_d\cap Y_i)\neq\emptyset \}$.  Observe that $F_{d}=F_{d_0}$ is contained in the interior of $F_{d_1}$, i.e., $N_{d_0}\subseteq N_{d_1}-F_{d_1}$.   Refer to \figref{d0d1}.  We claim that 
  \[  
     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \binom{k+1}{2} \enspace .
  \]
  \begin{figure}
    \begin{center}
      \includegraphics{figs/d0d1-5}
    \end{center}
    \caption{$F_{d_1},\ldots,F_{d_0}$ and the resulting graph $I$.}
    \figlabel{d0d1}
  \end{figure}
  To prove this claim, consider the graph $I$ with vertex set $V(I)=\{d_1,\ldots,d_0\}$ and $ij\in E(I)$ if and only $i=j$ or $|F_i\cap Y_j|\neq\emptyset$.  Any $\delta$-walk $v_0,\ldots,v_\ell$ with $\delta\in V(I)$ corresponds to a walk $\rho_0,\ldots,\rho_{\ell}$ in $I$, where $\rho_i=\max\{j\in\{0,\ldots,d\} : v_i\in N_{j}-F_j\}$.  Because of Properties~(i)--(iv) of $\delta$-walks, any walk $\rho_0,\ldots,\rho_\ell$ obtained this way has (i)~$\rho_0=d_0$, (ii--iii)~$\delta=\rho_\ell < \min\{\rho_0,\ldots,\rho_{\ell-1}\}$, and (iv)~$\ell \le k+1$.

  The graph $I$ has the following properties:
  \begin{enumerate}
    \item For each $i\in V(i)$, $id_1\in E(I)$.
    \item For each $i\in V(I)$, $i$ has at most 2 neighbours in $\{d_1+1,\ldots,i-1\}$. If $i$ has 2 such neighbours, then one of them is $i-1$.

    \item For each $ij\in E(I)$ with $j<i-1$, and each $i'\in\{j+1,\ldots,i+1\}$, the neighbours of $i'$ in $\{d_1+1,\ldots,i'-1\}$ are contained in $\{i'-1,j\}$.
  \end{enumerate}
  These two properties imply that $I$ has a very simple structure.   We call $j\in\{d_1,\ldots,d_0\}$ a \emph{left endpoint} if there exists an edge $ij\in E(I)$ with $j < i - 1$.  Let $j_1>j_2>\cdots> j_p$ denote the sorted sequence of left endpoints and, for convenience, let $j_{0}=d$.  Then 
  \[
      E(I) \subseteq \{ (i,i+1): i\in\{d_1,\ldots,d_0-1\} \cup 
         \bigcup_{t=1}^p \{ ij_t : i\in\{ j_t,\ldots,j_{t-1} \}\} \enspace . 
  \]
  (See the lower half of \figref{d0d1}.)  Observe that shortest path from in $i$ from $d_0$ to $j_t$ has length at least $t$.  Let $X_t=\{j_t,\ldots,j_{t+1}\}$ and observe that the shortest path from $d_0$ to some value $i\in X_t$ that does not contain any $i' <i$ goes through $j_{t}$.  Therefore, $X_t$ contains at most $k+2-t$ values of $\delta$ such that there exists a $\delta$-walk.  Therefore, 
  \[  
     |\{\delta\in\{d_1,\ldots,d_0-1\}:  \mbox{there exists a $\delta$-walk} \}| \le \sum_{t=0}^{k+1} (k+2-t) = \sum_{t=1}^{k+2} t = \binom{k+2}{2} \enspace .
  \]
